A parametric system for adaptive extrusion width contour parallel toolpath generation

### 2019-08-21 4_method.tex
big restructure based on new UoC overview

\subsection{Shape decomposition}
The shape decomposition we employ on the outline shape of a layer is based on a common skeletonization of the polygonal outline shape: the medial axis.
Starting from the medial axis we further decompose the shape into simple fragments, so that we can easily process the shape using piecewise linear approximations.

@@

\subsection{Bead counts}
We assign a bead count to each node in the skeleton which is in `the center' of the shape in a sense which will be defined in \cref{sec:significance_measure}.
Around regions where the bead count changes we introduce transition regions, so that we can smoothly change between regions with different bead count.

@@

@@

\subsection{Toolpath extraction}
\label{section_toolpath_extraction}
Now that we have decided on the bead counts in the central marked regions, we can apply those bead counts outward along the unmarked bones of the skeleton in order to define the locations of the toolpath junctions.
Because there is a plethora ways to cover a distance $d$ with $n$ beads, the placement of junctions depends on the chosen beading strategy.
The beading strategy is used to generate the bead widths and radial distances at which the junctions are placed for each node: the \emph{beading}.
That beading information is then propagated along the unmarked bones of the skeleton, so that we can generate the junctions on each unmarked bone.
The junctions are then connected together per trapezoid in order to form the final toolpaths.

\begin{definition}\label{beading_strategy_definition}
We define a beading strategy as the following set:
$$
\left\{ \alpha_\text{max}, b(d), t(n), t_0(n), W(n, d), L(n, d) \right\}
$$
where
$\alpha_{\text{max}}$ (the limit bisector angle),
$b(d)$ (optimal bead count),
$t(n)$ (transition length)
and
$t_0(n)$ (transition anchor position) have already been defined above.
We introduce here
$W(n, d)$, which gives the sequence of bead widths to cover a distance $d$ using $n$ beads
and
$L(n, d)$, which gives the sequence of toolpath positions as radial distances from the outline to cover a distance $d$ using $n$ beads.
\end{definition}

The following restrictions hold:
\begin{enumerate}
\item $W$ is symmetric: $W(n, d)_i = W(n, d)_{n-i-1}$
\item $L$ is symmetric in $d$: $L(n, d)_i = d - L(n, d)_{n-i-1}$
\item $W_n$ is monotonic and continuous at each bead index $n$ for constant bead count $c$: $0 \leq \frac{\partial W(c, d)_n}{\partial d} < \infty$.
\end{enumerate}

\begin{definition}
By applying the beading strategy to a skeleton location $v$ and a predetermined bead count $n$ we get a beading $B$, which is defined as the following set:
$$
\left\{ d, n, w_i, l_i  \right\}
$$
where
$d = 2 R(v)$ is the diameter associated with $v$,
$n$ is the bead count,
$w_i$ is the bead width of the bead with index $0 \leq i < n$ counting from the outline inward
and
$l_i$ is the distance of the toolpath location of bead with index $0 \leq i < n$.
\end{definition}

@@

\subsubsection{Toolpath extraction algorithm}
Now that each node has a beading associated with it we can apply the beading to generate the junctions of the toolpaths.
For each upward half-edge $e$ we generate each junction $J$ based on the beading $B = B^{v_1^e}$:
\begin{align*}
J &= \{ v, w, i \} \\ 
v &= v_1^e + (v_0^e - v_1^e) \frac{R(v_1^e) - l_i^B}{R(v_1^e) - R(v_0^e)} \\ 
w &= w_i^B
% \\ i^J &= i
\end{align*}
for any $i$ for which $R(v_0^e) < l_i^B \leq R(v_1^e)$.
See \cref{junction_placement}.
We store all junctions of a segment in a mapping from edge to a list of junctions.

>>>>

\subsection{Surface construction}
\todo{TODO: explain section title!}

The shape decomposition we employ on the outline shape of a layer is based on a common skeletonization of the polygonal outline shape: the medial axis.
Starting from the medial axis we further decompose the shape into simple fragments, so that we can easily process the shape using piecewise linear approximations.

@@

\subsection{Center classification}
\todo{TODO: explain section title!}

We assign a bead count to each node in the skeleton which is in `the center' of the shape in a sense which will be defined in \cref{sec:significance_measure}.
Around regions where the bead count changes we introduce transition regions, so that we can smoothly change between regions with different bead count.

@@

\subsection{Central height adjustment}
\todo{TODO: explain section title!}

\todo{Introduce what's to come in this section.}

@@

\subsection{Peripheral height adjustment}
\label{section_toolpath_extraction}

For our purposes we only need to define the height at the locations which will be part of a slice.
Instead of adjusting the height of peripheral nodes, we determine the locations along the bones where the adjusted height is an integer multiple of half the nozzle size.
This helps to define an efficient slicing algorithm in \cref{sec_toolpath_extraction}.

\todo{TODO:}
\begin{itemize}
\item explain section title
\item Reword and/or introduce concept of Beading 
\item Check compatibility with surrounding text!
\end{itemize}

Now that we have decided on the bead counts in the central marked regions, we can apply those bead counts outward along the unmarked bones of the skeleton in order to define the locations of the toolpath junctions.
Because there is a plethora ways to cover a distance $d$ with $n$ beads, the placement of junctions depends on the chosen beading strategy.
The beading strategy is used to generate the bead widths and radial distances at which the junctions are placed for each node: the \emph{beading}.
That beading information is then propagated along the unmarked bones of the skeleton, so that we can generate the junctions on each unmarked bone.
The junctions are then connected together per trapezoid in order to form the final toolpaths.

\todo{Definition of beading strategy got moved!}
\todo{Check compatibility with surrounding text!}

@@

\subsection{Toolpath extraction}\label{sec_toolpath_extraction}
Now that each node has a beading associated with it we can apply the beading to generate the junctions of the toolpaths.
For each upward half-edge $e$ we generate each junction $J$ based on the beading $B = B^{v_1^e}$:
\begin{align*}
J &= \{ v, w, i \} \\ 
v &= v_1^e + (v_0^e - v_1^e) \frac{R(v_1^e) - l_i^B}{R(v_1^e) - R(v_0^e)} \\ 
w &= w_i^B
% \\ i^J &= i
\end{align*}
for any $i$ for which $R(v_0^e) < l_i^B \leq R(v_1^e)$.
See \cref{junction_placement}.
We store all junctions of a segment in a mapping from edge to a list of junctions.

### 2019-08-21 5_generalization.tex
big restructure based on new UoC overview

\section{Beading strategies}
Now that we have described a framework for deciding and applying on bead counts and their widths locally, we can introduce several beading strategies which determine the bead count and their widths in different ways.
We can emulate a variety of toolpath generation methods from related literature by meticulously defining new beading strategies.
We also introduce new beading strategies which produce toolpaths with less extremal widths compared to techniques from existing literature.

@@

\section{Validation}
In order to test our framework, we have assembled a data set of different types of 3D models, ranging over various applications and various types of geometry.
The dataset is described in detail in the accompanying data in Appendix~\ref{dataset}.
We sliced all models in the dataset and selected 300 random slices from the total set of slices for thorough analysis.
We generated the toolpaths on these 300 outline shapes using the naive technique as implemented by Clipper and compare it to the constant bead count strategy with a bead count of $C=4$, the centered beading strategy, the evenly distributed, and the inward distributed beading strategy, using a nominal bead widths of $w_\text{pref}] = \SI{0.4}{\milli\meter}$.

\subsection{Computational analysis}
Experiments were performed on an Intel Core i7-7500U CPU @ \SI{2.70}{\giga\hertz} using a single core and \SI{16.3}{\giga\byte} memory.
The naive method was implemented using the state-of-the-art polygon offset library Clipper. \cite{johnson2014clipper}
The computation time of each slice from the dataset was recorded for the "Inward distributed" and "Naive" strategy. 
\cref{computime} shows these computation times of each layer plotted against the vertex count of the layer.

\paragraph{Accuracy}
In order to estimate the overfill and underfill, we need to accurately calculate the area covered by a single extrusion path.
If we would simply use an isosceles trapezoidal area with base lengths equal to the widths of the two end points of the segment, we would get strange artifacts at corners in the toolpath.
We therefore use a semi-circle with a diameter equal to the starting width in the one end of each segment, and exclude it at the other end, because it will be included in the next segment.
For polyline extrusion paths which are not closed, we also include the semi-circle corresponding to the extrusion width of the end-position.
In order to print such extrusion paths accurately, we can modulate the amount of material flow per millimeter based on this visualization model.
See \cref{segment_visualization}.

\begin{figure}
\centering
\setlength{\figwidth}{.25\columnwidth}
\begin{subfigure}{\figwidth}\centering
\includegraphics[width=\columnwidth]{sources/validation/visualization_principle_blocky.pdf}
\caption{Blocky}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[width=\columnwidth]{sources/validation/visualization_principle_rounded.pdf}
\caption{Rounded}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[width=\columnwidth]{sources/validation/visualization_principle_rounded_excluded.pdf}
\caption{Excluded}
\end{subfigure}
\caption{
Extruded area of two extrusion toolpath segments.
Red areas signify doubly extruded areas.
}
\label{segment_visualization}
\end{figure}

We can then estimate the amount of underfill by unioning the rounded visualization of all extrusion segments and then taking the difference from the original outline shape.
In order to deal with rounding errors, we perform a morphological close of \SI{5}{\micro\meter} before calculating the total area of the underfill regions.

The overfill regions are calculated by adding the polygonal areas of each extrusion segment to a list.
We then add the original outline shape in reverse and perform a clipper operation which only keeps areas of positive winding order.
Because the reverse outline shape reduces the winding order of all extruded segments by one, we are left with the overfill areas.
In order to estimate the areas which are triply covered by extrusion segments, we repeat the process of adding the outline in reverse and performing the clipper operation.
These triple extrusion areas count double toward the total amount of overfill.
The resulting overfill and underfill areas are visualized for the different toolpath strategies in the top of \cref{visualized_accuracy}.

\begin{figure*}
\centering
\setlength{\figwidth}{0.19\textwidth}
\setlength{\figheight}{0.283\textwidth}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_naive_accuracy.png}
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_naive_widths.png}
\caption{Naive}\label{TEST_naive_accuracy}
\end{subfigure}
%\begin{subfigure}{\figwidth}\centering
%\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_SingleBead_accuracy.png}
%\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_SingleBead_widths.png}
%\caption{Single}\label{TEST_SingleBead_accuracy}
%\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Constant_accuracy.png}
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Constant_widths.png}
\caption{Constant}\label{TEST_Constant_accuracy}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Center_accuracy.png}
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Center_widths.png}
\caption{Center}\label{TEST_Center_accuracy}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Distributed_accuracy.png}
\includegraphics[height=\figheight]{sources/validation/gMAT_example/TEST_Distributed_widths.png}
\caption{Distributed}\label{TEST_Distributed_accuracy}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[width=\columnwidth]{sources/validation/gMAT_example/TEST_InwardDistributed_accuracy.png}
\includegraphics[width=\columnwidth]{sources/validation/gMAT_example/TEST_InwardDistributed_widths.png}
\caption{Inward}\label{TEST_InwardDistributed_accuracy}
\end{subfigure}
\begin{subfigure}{.04\columnwidth}\centering
\vspace{4cm}
\includegraphics[height=\figheight]{sources/validation/gMAT_example/widths_legend.png}
\end{subfigure}
\caption{
Visualizing the overfills and underfills and the exaggerated widths for various beading strategies.
Grey areas are covered by extruded material,
blue areas in the top images signify underfill regions,
red areas overfill,
orange areas triply extruded areas (double overfill).
The bottom visualizes the bead widths at each location by exaggerating their widths.
Cold colors are narrower than the nozzle size and warm colors are wider than the nozzle size.
}
\label{visualized_accuracy}
\end{figure*}

We calculated the total overfill and underfill areas of the different toolpath strategies applied on the dataset. 
The overfill and underfill area as a percentage of the total area of the slices from the dataset are illustrated in \cref{over_underfill}.  
The "Inward distributed" strategy has a calculated overfill of 0.24\% and an underfill of 0.17\%.
This is lower compared to the "Naive" strategy, which results in 1.2\% overfill and 1.7\% underfill when applied to the dataset.

\begin{figure*}
\centering
\setlength{\figheight}{0.2\textwidth}
\setlength{\figwidth}{0.24\textwidth}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/computime2.pdf}
%\includegraphics[width=\figwidth]{sources/validation/computime2.pdf}
\caption{Computation time and target vertex count for the full dataset}
\label{computime}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/overunderfill.pdf}
%\includegraphics[width=\figwidth]{sources/validation/overunderfill.pdf}
\caption{Total over and underfill area}
\label{over_underfill}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/widthHistogram.pdf}
%\includegraphics[width=\figwidth]{sources/validation/widthHistogram.pdf}
\caption{Relative frequency of extrusion widths}
\label{widthHistogram}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheight]{sources/validation/smoothness.pdf}
%\includegraphics[width=\figwidth]{sources/validation/smoothness.pdf}
\caption{Frequency of angles between consecutive segments}
\label{smoothness}
\end{subfigure}
\caption{
Results of computational analysis of the results from applying the naive method and various beading strategies using our framework to the dataset.
Note the use of a logarithmic scale on the Y-axes and on some of the X-axes as well.
}
\end{figure*}

\paragraph{Uniformity}
Depending on the printer hardware configuration, some bead widths might be difficult or impossible to manufacture.
A bead width far below the nozzle size can result in fluttered extrusion, while a bead width far above the nozzle size can require pressures above the specifications of the machine. 
Also, large bead widths can cause the extruded material to bulge upwards instead of filling the required width.
\todo{Include picture of flutter?}
Moreover, bead widths deviating from the nominal bead width by a large amount may impact the mechanical properties of the print.
We visualize the bead widths resulting from the different strategies in an exaggerated manner in the bottom of \cref{visualized_accuracy}.

To evaluate the width uniformity, we sampled the extrusion widths of all generated toolpaths 
at intervals of \SI{200}{\micro\meter} along the toolpath.
From these sampled values, we calculated the mean bead width, and standard deviation.
\cref{widthHistogram} shows the distribution of extrusion width for each strategy binned at intervals of  \SI{10}{\micro\meter}.
The results show that the mean width of our strategies "Inwards distributed" and "Evenly distributed" is close to nominal nozzle size of \SI{400}{\micro\meter}, while their standard deviation is lower than for strategies "Centered" and "Constant bead count". 
These values indicate that, while causing less overfill and underfill, our strategy deviates less and less often from the nominal nozzle size.

\paragraph{Smoothness of toolpaths}
In order to maintain a high printing speed, it is desirable that toolpaths have fewer and less sharp corners. 
We measured the angle between consecutive segments generated by each strategy.
The relative frequency of these angles is shown in \cref{smoothness}.
All strategies show a higher number of corners below 45 degrees, with a peak towards 0 degrees.
While the strategy "constant bead width" has a lower number of corners overall, we found no significant differences in the smoothness of the toolpaths generated by the other strategies.

\subsection{Experimentation}
Test prints were performed on a custom FDM hardware setup, with a standard \SI{0.4}{\milli\meter} nozzle and a filament extrusion drive directly mounted on the print head.
The firmware of the printer employs linear advance: to gain the extra pressure required to change to a wider bead an extra amount of filament is advanced into the physical system between the filament feeder and the nozzle so that we can realize accurate adaptive deposition width control.\cite{tronvoll2019investigating}
We set the prefered width to $w_\text{pref} = \SI{0.6}{\milli\meter}$, so that we avoid fluttered printing of lines thinner than the nozzle size.
We used a layer thickness of \SI{0.2}{\milli\meter} and a movement speed of \SI{10}{\milli\meter\per\second}.

The resulting prints are shown in \cref{wedge_print} and \cref{prints}.
Because of inaccuracies in the depositioning control system some of the prints show defects.
Such defects are less prevalant for the inward distributed beading strategy than the other strategies.

\begin{figure}
\centering
\begin{subfigure}{\columnwidth}\centering
\setlength{\figwidth}{\columnwidth}
\includegraphics[width=\figwidth]{sources/applications/P3_print_wedge_naive_edited.png}
\includegraphics[width=\figwidth]{sources/applications/P3_print_wedge_center_edited.png}
\includegraphics[width=\figwidth]{sources/applications/P3_print_wedge_inward_edited.png}
\caption{Wedge}\label{print_wedge}
\end{subfigure}
\caption{
A wedge shape printed using the naive strategy, center deivation strategy and the inward distributed strategy.
The wedge print can be read like a graph with model diameter on the X-axis and the resulting beads over the Y-axis.
The naive and center deviation strategy show problems around certain model diameters.
}
\label{wedge_print}
\end{figure}

\begin{figure}
\centering
\setlength{\figheight}{.38\columnwidth}
\setlength{\figheightTwo}{.47\columnwidth}
\setlength{\figwidth}{0.32\columnwidth}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheightTwo]{sources/applications/gMAT_naive.png}
\includegraphics[height=\figheight]{sources/applications/P3_print_hex_naive_edited.png}
\includegraphics[width=\figwidth]{sources/applications/P3_print_UM_naive_edited.png}
\caption{Naive}\label{print_naive}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheightTwo]{sources/applications/gMAT_center.png}
\includegraphics[height=\figheight]{sources/applications/P3_print_hex_center_edited.png}
\includegraphics[width=\figwidth]{sources/applications/P3_print_UM_center_edited.png}
\caption{Center deviation}\label{print_center}
\end{subfigure}
\begin{subfigure}{\figwidth}\centering
\includegraphics[height=\figheightTwo]{sources/applications/gMAT_inward.png}
\includegraphics[height=\figheight]{sources/applications/P3_print_hex_inward_edited.png}
\includegraphics[width=\figwidth]{sources/applications/P3_print_UM_inward_edited.png}
\caption{Inward distributed}\label{print_inward}
\end{subfigure}
\caption{
Test shapes printed using the naive strategy, center deivation strategy and the inward distributed strategy.
The naive method produces distinct underfill areas.
The center deviation strategy shows some defects due to inaccurate control of extremal deposition widths.
The inward distributed strategy produces the least defects.
}
\label{prints}
\end{figure}

\iffalse
\todo{Maybe test the following qualities:}
\begin{itemize}
\item visual consistency of flat top skin surface
\item graphs of tensile tests on thin walled object
\end{itemize}
\fi

%We therefore visualize include a semi-circle with a diameter equal to the starting width in the one end, and exclude it at the other end, because it will be included in the next extrusion segment.

>>>>

\section{Generalization}
In this section we will generaize the described method into a parametric system, i.e. a geometric framework.

\todo{TODO:}
\begin{itemize}
\item introduce section
\end{itemize}

\subsection{Beading strategy}
Depending on the application, the hardware and the materials used we might want to have different strategies for filling a shape with contour parallel toolpaths.
we therefore..

\begin{definition}\label{beading_strategy_definition}
We define a beading strategy as the following set:
$$
\left\{ \alpha_\text{max}, b(d), t(n), t_0(n), W(n, d), L(n, d) \right\}
$$
where
$\alpha_{\text{max}}$ (the limit bisector angle),
$b(d)$ (optimal bead count),
$t(n)$ (transition length)
and
$t_0(n)$ (transition anchor position) have already been defined above.
We introduce here
$W(n, d)$, which gives the sequence of bead widths to cover a distance $d$ using $n$ beads
and
$L(n, d)$, which gives the sequence of toolpath positions as radial distances from the outline to cover a distance $d$ using $n$ beads.
\end{definition}

The following restrictions hold:
\begin{enumerate}
\item $W$ is symmetric: $W(n, d)_i = W(n, d)_{n-i-1}$
\item $L$ is symmetric in $d$: $L(n, d)_i = d - L(n, d)_{n-i-1}$
\item $W_n$ is monotonic and continuous at each bead index $n$ for constant bead count $c$: $0 \leq \frac{\partial W(c, d)_n}{\partial d} < \infty$.
\end{enumerate}

\begin{definition}
By applying the beading strategy to a skeleton location $v$ and a predetermined bead count $n$ we get a beading $B$, which is defined as the following set:
$$
\left\{ d, n, w_i, l_i  \right\}
$$
where
$d = 2 R(v)$ is the diameter associated with $v$,
$n$ is the bead count,
$w_i$ is the bead width of the bead with index $0 \leq i < n$ counting from the outline inward
and
$l_i$ is the distance of the toolpath location of bead with index $0 \leq i < n$.
\end{definition}

\subsection{Beading strategies}
\todo{This section was moved. Check compatibility with surrounding text.}

Now that we have described a framework for deciding and applying on bead counts and their widths locally, we can introduce several beading strategies which determine the bead count and their widths in different ways.
We can emulate a variety of toolpath generation methods from related literature by meticulously defining new beading strategies.
We also introduce new beading strategies which produce toolpaths with less extremal widths compared to techniques from existing literature.

@@
